Navigating to a greener Europe with 24/7 hourly carbon-free electricity procurement

### 2023-10-11 manuscript/sections/introduction.tex
manuscript: polishing sections 1-3

Many public and private energy buyers join this effort by purchasing clean energy to demonstrate their sustainability credentials. The established schemes for voluntary clean energy procurement are the Guarantees of Origin (\gls{go}s) in Europe and the Renewable Energy Certificates (also known as Renewable Energy Credits) (\gls{rec}s) in the United States.
A common feature of these schemes is that renewable energy credits are \enquote{unbundled} from megawatt-hours of energy and can be traded numerous times on specialized markets before being retired.
The owner gets to claim environmental benefits when \enquote{retiring} (i.e. using) the certificate.
The unbundled renewable energy credits, however, have been criticized for not ensuring a physical match between renewable energy generation and consumption, and for absence of evidence that such schemes drive build-out of additional renewable energy capacity \cite{bjorn-RECSnatcom-2022, gillenwater-2014, re100report-2020, spglobal-recs, bock-icelandGOproblem}.

@@

\textbf{Existing literature on 24/7 CFE --} The growing interest in the means, costs and the system-level impacts of hourly \gls{cfe} procurement created a need for a quantitative analysis of the approach.
The existing model-based studies on 24/7~CFE approach are very recent.
The first to incorporate it into the electricity system planning model were Xu et al. (2022) \cite{xu-247CFE-SSRN}.\footnote{A working paper updates a prior public technical report published in November 2021 investigating the system-level impacts of 24/7~CFE in two regions of the U.S. -- California and PJM \cite{xu-247CFE-report}.} 
The authors study the system-level impacts of 24/7~CFE procurement based on an example of California in the year 2030.
In November 2022, IEA also released a model-based study on the 24/7~CFE, focusing on India and Indonesia \cite{ieaAdvancingDecarbonisationClean2022}.
In early 2023, Peninsula Clean Energy published an own white paper introducing their modeling tool---MATCH (Matching Around-the-Clock Hourly energy)---and the results of the modeling indicating feasibility of the company's \enquote{24/7 renewable energy by 2025} goal \cite{peninsula-report247}.

@@

\textbf{Contribution --} The goal of this work is to systematically explore different designs, optimal procurement strategies, costs, benefits and impacts of the 24/7~CFE matching approach, both for participating buyers and for regions where voluntary procurement occurs.
Our analysis focuses on four regions within the European energy system with unique characteristics, and we vary the analysis years to track the evolution of the energy system.
Additionally, we examine the underlying mechanisms that drive the system-level reduction of \co emissions, and examine how they perform over time and across regions.

>>>>

Many public and private energy buyers join this effort by purchasing clean energy to demonstrate their sustainability credentials. The established schemes for voluntary clean energy procurement are the Guarantees of Origin (\gls{go}s) in Europe and the Renewable Energy Certificates (\gls{rec}s) in the United States.
A common feature of these schemes is that renewable energy credits are \enquote{unbundled} from megawatt-hours of energy and can be traded numerous times on specialized markets before being retired.
The owner gets to claim environmental benefits when \enquote{retiring} (i.e. using) the certificate.
The unbundled renewable energy credits, however, have been criticized for not ensuring a physical match between renewable energy generation and consumption, and for absence of evidence that such schemes drive build-out of additional renewable energy capacity \cite{bjorn-RECSnatcom-2022, gillenwater-2014, re100report-2020, spglobal-recs, bock-icelandGOproblem}.

@@

\textbf{Existing literature on 24/7 CFE --} The growing interest in the means, costs and the system-level impacts of hourly \gls{cfe} procurement created a need for a quantitative analysis of the approach.
The existing model-based studies on 24/7~CFE approach are very recent.
The first to incorporate it into the electricity system planning model were Xu et al. (2022) \cite{xu-247CFE-SSRN}.\footnote{A working paper updates a prior report published in November 2021 investigating the system-level impacts of 24/7~CFE in two regions of the U.S. -- California and PJM \cite{xu-247CFE-report}.} 
The authors study the system-level impacts of 24/7~CFE procurement based on an example of California in the year 2030.
In November 2022, IEA also released a model-based study on the 24/7~CFE, focusing on India and Indonesia \cite{ieaAdvancingDecarbonisationClean2022}.
In early 2023, Peninsula Clean Energy published an own white paper introducing their modeling tool---MATCH (Matching Around-the-Clock Hourly energy)---and the results of the modeling indicating feasibility of the company's \enquote{24/7 renewable energy by 2025} goal \cite{peninsula-report247}.

@@

\textbf{Contribution --} The goal of this work is to systematically explore different designs, optimal procurement strategies, costs, benefits and impacts of the 24/7~CFE matching approach, both for participating buyers and for regions where voluntary procurement occurs.
Our analysis focuses on four regions within the European energy system with unique characteristics, and we vary the analysis years to track the evolution of the energy system.
Additionally, we examine the underlying mechanisms that drive the system-level reduction of \co emissions, and examine how they perform over time and across regions. 
This work is a follow-up to the earlier public report by the authors \cite{riepin-zenodo-systemlevel247} with an updated model, updated system-wide assumptions, and additional analysis of the results.

### 2023-10-11 manuscript/sections/methods.tex
manuscript: polishing sections 1-3

\textbf{Optimisation model --} The simulations were carried out with the modified version of the PyPSA-Eur model \cite{PyPSA-Eur-github}. 
The objective of the model is to minimise the annualised system-wide costs of meeting energy demand within the modelled system, while adhering to all relevant engineering, reliability, and policy constraints.
The optimal solution of the model includes the location and capacity of new power generation and storage assets, as well as the hourly dispatch of all existing and new assets, including the linear optimal power flow on the transmission network for the planning year. This is standard for the PyPSA-Eur model \cite{horschPyPSAEurOpenOptimisation2018} and the long-term energy planning models of this class \cite{jenkinsGenX2022, howellsOSeMOSYSOpenSource2011}.

\textbf{Clean energy procurement --} We introduce new model components (parameters, variables, constraints) to model voluntary clean energy procurement.
By incorporating these into the mathematical problem, we model a situation when a fraction of electricity demand in a particular bidding zone is committed to procuring additional clean energy with a desired strategy.
The fraction is set at 10\% of the total \gls{ci} electricity demand in a particular bidding zone, unless otherwise stated (hereinafter referred to as \enquote{participating consumers}).
Accordingly, the model co-optimises the procurement and operational decisions of the participating consumers to meet their electricity demand in accordance with the procurement strategy they have chosen.
The mathematical formulation of the clean energy procurement strategies is provided in the supplementary material [\labelcref{sec:si_3}].

The 24/7~CFE procurement framework is based on the methodologies and metrics paper by Google (2021) \cite{google-methodologies}.
The framework considers various \gls{cfe} supply options for the participating consumers, including procurement of own generation resources and storage assets, and electricity imports from the local grid.
The framework also allows for controlling \textit{the quality score} of the 24/7~CFE procurement by setting the threshold for the minimum share of hours when the hourly \gls{cfe} supply meets the hourly electricity demand.

@@

\begin{itemize}[-]
    \item \textit{100\% annual matching}: participating buyers meet their annual energy demand on volumetric basis with additional renewable energy procurement;
    \item \textit{24/7 hourly matching}: participating buyers optimise their investment and operational decisions mixing procurement of additional generation and storage resources and electricity imports so that \gls{cfe} supply meets electricity demand 24/7 throughout the year with the desired quality score;
    \item For benchmarking, we also model \textit{a reference case}: participating buyers have no voluntary environmental commitments and get all their electricity from the local grid.
\end{itemize}

We setup the study design so that the procurement strategies are modelled under the same conditions.
In both cases (i) procurement is possible with generators from the same bidding zone only (i.e., we model the \enquote{best case} of the annual matching strategy); (ii) the same set of consumers participate; (iii) the same set of constraints is applied to the rest of the electricity system.

>>>>

\textbf{Optimisation model --} The simulations were carried out with the modified version of the PyPSA-Eur model \cite{PyPSA-Eur-github}. 
The objective of the model is to minimise the annualised system-wide costs of meeting energy demand within the modelled system, while adhering to all relevant engineering, reliability, and policy constraints.
We use a brownfield investment approach, which means that we keep the existing power generation and storage assets in the model and only optimise the capacity expansion of new assets.
The optimal solution of the model includes the location and capacity of new power generation and storage assets, as well as the hourly dispatch of all existing and new assets, including the linear optimal power flow on the transmission network for the planning year. 
This is standard for the PyPSA-Eur model \cite{horschPyPSAEurOpenOptimisation2018} and the long-term energy planning models of this class \cite{jenkinsGenX2022, howellsOSeMOSYSOpenSource2011}.

\textbf{Clean energy procurement --} We introduce new model components (parameters, variables, constraints) to model voluntary clean energy procurement.
By incorporating these into the mathematical problem, we model a situation when a fraction of electricity demand in a particular bidding zone is committed to procuring additional clean energy with a desired strategy.
The fraction is set at 10\% of the total \gls{ci} electricity demand in a particular bidding zone, hereinafter referred to as \enquote{participating consumers}.
The participating consumers are modelled as a single entity with a baseload demand profile.
Accordingly, the model co-optimises the procurement and operational decisions of the participating consumers to meet their electricity demand in accordance with the procurement strategy they have chosen.
The mathematical formulation of the clean energy procurement strategies is provided in the supplementary material [\labelcref{sec:si_3}].

The 24/7~CFE procurement framework is based on the methodologies and metrics paper by Google (2021) \cite{google-methodologies}.
The framework considers various \gls{cfe} supply options for the participating consumers, including procurement of own generation resources and storage assets, and electricity imports from the local grid.
The framework also allows for controlling \textit{the quality score} of the 24/7~CFE procurement by setting the target threshold for the minimum share of hours when the hourly \gls{cfe} supply meets the hourly electricity demand.
The quality score is hereinafter referred to as \enquote{CFE target}, which takes values in percent points, e.g., 95\%, 99\%, 100\%.
The higher the CFE target, the higher the quality of the 24/7~CFE supply; see equation [TODO] for details.

@@

\begin{itemize}[-]
    \item \textit{100\% annual matching}: participating buyers meet their annual energy demand on volumetric basis with additional renewable energy procurement;
    \item \textit{24/7 hourly matching}: participating buyers optimise their investment and operational decisions mixing procurement of additional generation and storage resources and electricity imports so that \gls{cfe} supply meets electricity demand 24/7 throughout the year with the desired quality target.
\end{itemize}

We setup the study design so that the procurement strategies are modelled under the same conditions.
In both cases (i) procurement is possible with generators from the same bidding zone only (i.e., we model the \enquote{best case} of the annual matching strategy); (ii) the same set of consumers participate; (iii) the same set of constraints is applied to the rest of the electricity system. 
For benchmarking, we also model \textit{a reference case}: participating buyers have no voluntary environmental commitments and purchase all their electricity from the local grid.

### 2023-10-11 manuscript/sections/results.tex
manuscript: polishing sections 1-3

Our experimental setup incorporates a wide range of scenarios, allowing us to generalise our results beyond individual assumptions.
For convenience, we start with a \textit{base scenario} presented in \cref{subsec:base}.
The base scenario reflects a case when 10\% (on volumetric basis) of \gls{ci} consumers in \textit{Ireland} in \textit{2025} are committed to procuring additional clean energy with a desired strategy.
The participating consumers can attain the 24/7~CFE procurement targets by co-optimizing local procurement using mature technologies, designated as \textit{palette~1}, and by importing electricity from the local grid.

Afterward, we delve deeper into the scenario space, altering highlighted parameters to explore various alternatives and address the following questions:

@@

Considering the large scenario space, this work presents results for a selection of scenario branches that generalise the system impacts of 24/7 CFE procurements.
For interested readers, we publish an online dashboard with results for the entire scenario space, see \nameref{sec:code}.

@@

%some parts of this paragraph may go to methods
\cref{fig:10-2025-IE-p1-used} shows the fraction of the fraction of hourly demand met with carbon-free electricity depending on the procurement policy that participating consumers follow.
We contrast two main cases, one where participating consumers commit to 100\% annual renewable matching policy (hereinafter -- \enquote{100\% RES}) and another case where the consumers commit to 24/7~CFE hourly matching policy (hereinafter -- \enquote{24/7~CFE}) with various \gls{cfe} targets in a range from 80\% to 100\%. 
In the \textit{reference case}, participating consumers cover their load purely with grid purchases without any policy regarding the origin of electricity.

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.95\columnwidth}
        \centering
        \caption{Average emissions rate of participating consumers.}
    \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/ci_emisrate.pdf}
    \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/ci_emisrate.pdf}
    \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/ci_emisrate.pdf}
    \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/ci_emisrate.pdf}
    \caption{Average emissions rate of participating consumers.}
        \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/ci_emisrate.pdf}
    \caption{Average emissions rate of participating consumers.}
        \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/ci_emisrate.pdf}
    \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/ci_emisrate.pdf}
    \caption{Average emissions rate of participating consumers.}
        \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/ci_emisrate.pdf}
    \caption{Average emissions rate of participating consumers.}
        \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/ci_emisrate.pdf}
    \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/ci_emisrate.pdf}
    \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/ci_emisrate.pdf}
    \caption{Average emissions rate of participating consumers.}
        \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/ci_emisrate.pdf}
    \caption{Average emissions rate of participating consumers.}
        \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/ci_emisrate.pdf}
    \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/ci_emisrate.pdf}
    \caption{Average emissions rate of participating consumers.}
        \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/ci_emisrate.pdf}
    \caption{Average emissions rate of participating consumers.}
        \label{fig:10-2025-IE-p1-ci_emisrate}
    \end{subfigure}
    \begin{subfigure}[t]{0.95\columnwidth}
        \centering
        \vspace{-0.2cm}
        \caption{\co emissions in the local region (Ireland).}
    \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/zone_emissions.pdf}
    \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/zone_emissions.pdf}
    \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/zone_emissions.pdf}
    \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/zone_emissions.pdf}
    \caption{\co emissions in the local region (Ireland).}
        \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/zone_emissions.pdf}
    \caption{\co emissions in the local region (Ireland).}
        \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/zone_emissions.pdf}
    \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/zone_emissions.pdf}
    \caption{\co emissions in the local region (Ireland).}
        \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/zone_emissions.pdf}
    \caption{\co emissions in the local region (Ireland).}
        \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/zone_emissions.pdf}
    \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/zone_emissions.pdf}
    \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/zone_emissions.pdf}
    \caption{\co emissions in the local region (Ireland).}
        \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/zone_emissions.pdf}
    \caption{\co emissions in the local region (Ireland).}
        \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/zone_emissions.pdf}
    \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/zone_emissions.pdf}
    \caption{\co emissions in the local region (Ireland).}
        \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/zone_emissions.pdf}
    \caption{\co emissions in the local region (Ireland).}
        \label{fig:10-2025-IE-p1-zone_emissions}
    \end{subfigure}
    \caption{Impact of clean electricity procurement commitments on emissions of participating consumers (top panel) and the system-level emissions in a local region (bottom panel).}
    \label{fig:10-2025-IE-p1-emissions}

>>>>

Our experimental setup incorporates a wide range of scenarios, allowing us to generalise our results beyond individual assumptions.
For convenience, we start with a \textit{base scenario} presented in \cref{subsec:base}.
The base scenario reflects a case when 10\% (on volumetric basis) of \gls{ci} consumers in \textit{Ireland} in \textit{2025} are committed to procuring additional clean energy with a desired strategy.
The participating consumers can attain the 24/7~CFE procurement targets by co-optimizing local procurement using mature technologies, designated as \textit{palette~1}, and by importing electricity from the local grid. Afterward, we delve deeper into the scenario space, altering highlighted parameters to explore various alternatives and address the following questions:

@@

Considering the large scenario space, this work presents results for a selection of scenario branches that generalise the system impacts of 24/7 CFE procurements.
We publish an online dashboard with results for the entire scenario space alongside the paper, see \nameref{sec:code}.

@@

%some parts of this paragraph may go to methods
\cref{fig:10-2025-IE-p1-used} shows the fraction of hourly demand met with \gls{cfe} depending on the procurement policy that participating consumers follow.
We contrast two main cases, one where participants commit to 100\% annual renewable matching policy (hereinafter -- \enquote{100\% RES}) and another case where the consumers commit to 24/7~CFE hourly matching policy (hereinafter -- \enquote{24/7~CFE}) with various \gls{cfe} targets in a range from 80\% to 100\%. 
In the \textit{reference case}, participating consumers cover their load purely with grid purchases without any policy regarding the origin of electricity.

@@

\begin{figure}
    \centering
    \begin{subfigure}[t]{0.95\columnwidth}
        \centering
        \caption{Average emissions rate of participating consumers.}
        \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/ci_emisrate.pdf}
        \label{fig:10-2025-IE-p1-ci_emisrate}
    \end{subfigure}
    \begin{subfigure}[t]{0.95\columnwidth}
        \centering
        \vspace{-0.2cm}
        \caption{\co emissions in the local region (Ireland).}
        \includegraphics[width=0.95\columnwidth]{plots/10/2025/IE/p1/zone_emissions.pdf}
        \label{fig:10-2025-IE-p1-zone_emissions}
    \end{subfigure}
    \caption{Impact of clean electricity procurement commitments on emissions of participating consumers (top panel) and the system-level emissions in a local region (bottom panel).}
    \label{fig:10-2025-IE-p1-emissions}
\end{figure}
